Neural Connectome of the Ctenophore Statocyst

### 2025-04-30 manuscript/Manuscript.qmd
Fig4

We found that arrest and re-beat events were synchronized between balancers across the sagittal plane. However, while the timing of re-beat was also near-simultaneous along the tentacular plane, arrest timing was offset by up to 2.xy <!--# add max value--> seconds (Figure 4) <!--# save source data for all plots in manuscript/source_data with the naming convention e.g. Figure 1A--source data 1 -->. Overall, these data reveal that arrests are only coordinated between Q1-Q2 and Q3-Q4, whereas re-beat is coordinated over all four quadrants.

<!--# discuss frequency data -->

@@

![**Figure 4. Differences in Balancer Cilia Movement Control Due to Variations in the Neural Circuit Pathways**\
(A) Schematic diagram of differential interference contrast (DIC) microscopy setup for imaging balancer cilia movement. The microscope was tilted 90 degrees so that the stage was positioned vertically. A monochrome CMOS camera sensitive to near-infrared (NIR) light was used, synchronized with an 850 nm strobe light source. The movement of the balancer cilia was recorded using a 40× objective lens.\
(B) Kymograph patterns comparing the arrest and re-beat of balancer cilia movement. The left and right balancer cilia movements are compared in both the sagittal plane (left) and tentacular plane (right). The direction and length of the arrows indicate time.\
(C) Graph showing the time differences in the arrest and re-beat timing of the left and right balancer cilia. S and T represent the sagittal plane and tentacular plane, respectively.](figures/Figure4.png)

>>>>

We found that arrest and re-beat events were synchronized between balancers across the sagittal plane. However, while the timing of re-beat was also near-simultaneous along the tentacular plane, arrest timing was offset by up to 2.16 seconds (Figure 4D). Overall, these data reveal that arrests are only coordinated between Q1-Q2 and Q3-Q4, whereas re-beat is coordinated over all four quadrants.

As shown in Figure 4E, beat frequencies of left and right balancers fluctuated substantially in both orientations. However, while left-right activity was largely synchronized in sagittal pairs (Figure 4E, left), tentacular pairs (Figure 4E, right) showed strikingly periodic alternations between left and right balancer activity. This pattern suggested an offset in timing during arrest or re-beat events between tentacular balancers.

To quantify this, we calculated a rolling Pearson correlation between left and right balancer activity in each plane. The resulting mean correlations were significantly higher in the tentacular plane compared to the sagittal plane (Figure 4F; t-test, p = 0.014). This result indicates that tentacular balancers beat in a more coordinated, mirror-symmetric manner, whereas sagittal balancers exhibit greater variability between sides.

@@

Notably, bridge neurons associated with each quadrant region (Q1Q2 or Q3Q4) extend their processes across the midline, forming connections that span between opposite quadrant domains when viewed along the tentacular plane. This morphology suggests a potential role for bridge neurons in modulating inter-quadrant coordination. By linking ANN activity across orthogonal planes, these neurons may contribute to the dynamic regulation of CBF patterns across the whole system.

![**Figure 4. Differences in Balancer Cilia Movement Control Due to Variations in the Neural Circuit Pathways**\
(A) Schematic diagram of differential interference contrast (DIC) microscopy setup for imaging balancer cilia movement. The microscope was tilted 90 degrees so that the stage was positioned vertically. A monochrome CMOS camera sensitive to near-infrared (NIR) light was used, synchronized with an 850 nm strobe light source. The movement of the balancer cilia was recorded using a 40× objective lens.\
(B) Representative DIC images of the aboral organ viewed along the sagittal (left) and tentacular (right) planes. Arrowheads indicate the balancer cilia selected for kymograph-based analysis in each orientation.\
(C) Representative kymographs showing arrest and re-beat events of balancer cilia. Left and right balancer cilia were simultaneously recorded in either the sagittal (left) or tentacular plane (right). Arrowheads mark the onset of arrest (open) and re-beat (filled); horizontal arrows represent 1-second intervals.\
(D) Boxplots showing time differences in the onset of arrest (white boxes) and re-beat (gray boxes) between left and right balancer cilia. Colors of the box outlines indicate the imaging plane: sagittal (blue) or tentacular (orange). Each dot represents a single larva.\
(E) Bar plots showing ciliary beat frequency (CBF) dynamics of left and right balancers (top and bottom, respectively) recorded from sagittal (left) and tentacular (right) planes. Positive and negative values represent opposite sides within each plane.\
(F) Mean rolling Pearson correlation of left–right balancer CBF calculated with a 20-frame window. Each dot represents an individual larva. Balancers on the tentacular plane showed significantly higher correlation than those on the sagittal plane (t-test, p = 0.014).](figures/Figure4.png)
